---
title: "fig2"
author: "Daniel Chawla"
date: "3/18/2022"
output: md_document
---

```{r, include = FALSE, echo = FALSE}
data_list <- readRDS('~/Documents/darpa-manuscript-data/data_list_final/data_list_v1.RDS')
```


```{r, include = FALSE}
library(dplyr)
library(ggplot2)
library(ggpubr)
library(stringr)
library(patchwork)
library(ggpattern)

source('../analysis/plot_palette.R')

```


# Figure 2C-E
Here, we characterize technical factors in viral and bacterial datasets that may bias downstream evaluation.

Note: final visual styling for the manuscript was done in post-processing

Figure 2C: study sizes (number of samples)

Figure 2D: time series vs. cross-sectional study designs 

Figure 2E: platform manufacturers 


## Figure 2C: Study sizes for viral vs. bacterial datasets
```{r}
# function that takes in a MetaIntegrator object MIobj, 
# positive and negative class information, and a minimum sample size (default 4)
# and returns an index indicating whether the dataset should be included (T) or
# excluded (F)
findStudyWithClass <- function(MIobj, pos_class, neg_class, N_threshold = 4){
  # logical indices for samples in the positive and negative classes
  pos_ix <- sum(MIobj$pheno$Class %in% pos_class)
  neg_ix <- sum(MIobj$pheno$Class %in% neg_class)
  
  # T if both positive and negative classes are above the minimum sample size 
  if(sum(pos_ix) >= N_threshold & sum(neg_ix) >= N_threshold){
    return(T)
  } else {
    return(F)
  }
}

# find bacterial studies
bac_ix <- sapply(data_list, 
                 findStudyWithClass, 
                 pos_class = 'Bacteria', 
                 neg_class = c('Healthy', 'Other.Infectious', 'Other.NonInfectious', 'Convalescent'))

## print bacterial class vectors to sanity check that correct studies were identified
# sapply(data_list[bac_ix], function(x){table(x$pheno$Class)})

# find viral studies
vir_ix <- sapply(data_list, 
                 findStudyWithClass, 
                 pos_class = 'Virus', 
                 neg_class = c('Healthy', 'Other.Infectious', 'Other.NonInfectious', 'Convalescent'))

## print viral class vectors to sanity check that correct studies were identified
#sapply(data_list[vir_ix], function(x){table(x$pheno$Class)})

# function that takes a MetaIntegrator object (MIobj) input and a list of 
# desired classes within $pheno and returns a sample size
getNClass <- function(MIobj, input_classes){
  return(sum(MIobj$pheno$Class %in% input_classes))
}

# bacterial and viral sample sizes
bac_N <- sapply(data_list[bac_ix], getNClass, input_classes = c('Bacteria', 'Healthy', 'Other.Infectious', 'Other.NonInfectious'))
vir_N <- sapply(data_list[vir_ix], getNClass, input_classes = c('Virus', 'Healthy', 'Other.Infectious', 'Other.NonInfectious'))

# format for plotting
hist_plot_dat <- data.frame(
  'Type' = c(rep('Virus', length(vir_N)),
             rep('Bacteria', length(bac_N))),
  'N' = c(vir_N, bac_N))
hist_plot_dat <- hist_plot_dat %>% 
  mutate(Type = factor(Type, levels = c('Virus', 'Bacteria'), labels = c('viral', 'bacterial'), ordered = T))

# plots of sample sizes
## histogram
#ggplot(hist_plot_dat, aes(x = N, fill = Type)) + geom_histogram()
## boxplots
color_vals <- c('#C4E8F0', '#F2BE84')
names(color_vals) <- c('viral', 'bacterial')

p2c <- ggplot(hist_plot_dat, aes(x = Type, y = N, fill = Type)) + 
  geom_boxplot() + 
  scale_fill_manual(values = color_vals) + 
  stat_compare_means(method = 'wilcox.test') + 
  theme_clean
p2c
```

## Time-series vs. cross-sectional study designs for viral vs. bacterial datasets
```{r}
# function
## input: MetaIntegrator object (MIobj)
## output: logical indicating time series (T) or cross-sectional (F)
findTS <- function(MIobj){
  # MIobjs with $pheno containing $Time.Point were annotated as time series
  if(sum(names(MIobj$pheno) == 'Time.Point') > 0){
    return(T)
  } else{
    return(F)
  }
}

# ID time-series datasets
time_series_ix <- sapply(data_list, findTS)
ts_list <- data_list[time_series_ix]

# count time-series viral and bacterial datasets
ts_bac_ix <- sapply(ts_list, findStudyWithClass, pos_class = 'Bacteria', neg_class = c('Healthy', 'Other.NonInfectious', 'Convalescent'))
ts_vir_ix <- sapply(ts_list, findStudyWithClass, pos_class = 'Virus', neg_class = c('Healthy', 'Other.NonInfectious', 'Convalescent'))
ts_non_ix <- sapply(ts_list, findStudyWithClass, pos_class = 'Other.Infectious', neg_class = c('Healthy', 'Other.NonInfectious', 'Convalescent'))

## Counting and sanity checks
#sum(ts_bac_ix)
#sum(ts_vir_ix)
#sum(ts_non_ix)

#sapply(ts_list[ts_non_ix], function(x){table(x$pheno$Pathogen)})
## GSE28405_GPL2700 is viral
## GSE68310_GPL10558 is viral
## GSE7000_GPL4857 is bacterial and other.infectious
## GSE35859_GPL15241 is other.infectious
#sapply(ts_list[ts_vir_ix], function(x){table(x$pheno$Pathogen)})
## these are all correct
# sapply(ts_list[ts_bac_ix], function(x){table(x$pheno$Pathogen)})

# get time-series counts per pathogen type
N_other <- 2 # 7000 is common with bacteria (manually checked)
N_vir <- sum(ts_vir_ix) # all look correct
N_bac <- sum(ts_bac_ix) # 2 are in common with the virus list (40012 and 20346)

# get cross-sectional counts per pathogen type
vir_accs <- names(data_list[vir_ix]) # snapshot studies are the unused from each category
bac_accs <- names(data_list[bac_ix])
if(mean(names(ts_list[ts_vir_ix]) %in% vir_accs) != 1){
  stop('not all viral ts in full ts list')
}
N_vir_snapshot <- sum(!vir_accs %in% names(ts_list[ts_vir_ix]))
N_bac_snapshot <- sum(!bac_accs %in% names(ts_list[ts_bac_ix]))

# format plot data
ts_N_df <- data.frame(Infection = c('Virus', 'Virus', 'Bacteria', 'Bacteria'), 
                      Type = c('Time Series', 'Snapshot', 'Time Series', 'Snapshot'),
                      N = c(N_vir, N_vir_snapshot, N_bac, N_bac_snapshot)) %>% 
  mutate(Infection = factor(Infection, levels = c('Virus', 'Bacteria'), labels = c('viral', 'bacterial'), ordered = T))%>%
  mutate(Type = factor(Type, levels = c('Time Series', 'Snapshot'), ordered = T))

# plot
color_vals <- c('#C4E8F0', '#F2BE84')
names(color_vals) <- c('viral', 'bacterial')

# plot
p2d <- ts_N_df %>%
  ggplot(aes(x = Infection, y = N, fill = Infection, pattern = Type)) + 
  geom_bar_pattern(stat = 'identity') + 
  theme_clean + 
  scale_fill_manual(values = color_vals) + 
  labs(x = 'Study Design', y = 'Number of Studies')
p2d
```

## Platform manufacturer for viral and bacterial datasets
```{r}
# function
## input: dataset name
## output: GEO platform accession as GPL####
parsePlatform <- function(str){
  str_extract(pattern = 'GPL\\d+', string = str) %>%
    return()
}

# function
## input: list of MetaIntegrator objects (input_list), pathogen type str (type)
## output: data frame, sorted count of platforms with pathogen type labeled
getPlatformsFromList <- function(input_list, type){
  input_list %>%
    names() %>%
    parsePlatform() %>%
    table() %>%
    sort(decreasing = T) %>%
    as.data.frame() %>%
    mutate('Type' = type) %>%
    return()
}

# get viral and bacterial platform counts
vir_platforms <- getPlatformsFromList(data_list[vir_ix], 'Virus')
bac_platforms <- getPlatformsFromList(data_list[bac_ix], 'Bacteria')

# format for plotting
platform_df <- bind_rows(vir_platforms, bac_platforms)
names(platform_df)[1] <- 'Platform'

# manually encode main platforms found in data that are not listed in the master table
appending_platforms <- 
data.frame('Platform' = c('GPL6254',   'GPL14604', 'GPL15615', 'GPL17586', 'GPL17692', 'GPL21947', 'GPL23126', 'GPL8300', 'GPL9392', 'GPL13287', 'GPL16951', 'GPL20844', 'GPL6254'),
       'Manufacturer' = c('other',     'affy',     'other',    'affy',     'affy',     'other',    'affy',     'affy',    'affy',    'other',    'other',    'other',    'other'))

# load main platform annotation table linking GPL IDs to manufacturers
annotation_df <- openxlsx::read.xlsx(paste0(Sys.getenv('DARPA'), 'Input/platform_specific/platform_types.xlsx')) %>%
  filter(!is.na(Manufacturer)) %>% 
  bind_rows(appending_platforms)

# add manufacturers to list of data platforms
platform_df <- platform_df %>%
  left_join(annotation_df)

# filter down to Illumina and Affymetrix, label everything else "other"
manufacturer_df <- platform_df %>%
  mutate(Manufacturer = ifelse(!Manufacturer %in% c('ilmn', 'affy'), 'other', Manufacturer)) %>% 
  group_by(Manufacturer, Type) %>%
  summarize('Freq' = sum(Freq)) %>%
  mutate(Manufacturer = factor(Manufacturer, 
                               levels = rev(c('ilmn', 'affy', 'other')), 
                               labels = rev(c('Illumina', 'Affymetrix', 'Other')), 
                               ordered = T)) %>%
  mutate(Type = factor(Type, levels = c('Virus', 'Bacteria'), labels = c('viral', 'bacterial'), ordered = T))

# plot
color_vals <- c('#C4E8F0', '#F2BE84')
names(color_vals) <- c('viral', 'bacterial')

p2e <- ggplot(manufacturer_df, aes(x = Type, y = Freq, fill = Type)) + 
  geom_bar_pattern(aes(pattern = Manufacturer), stat = 'identity') + 
  #geom_bar(stat = 'identity') + 
  theme_clean + 
  scale_fill_manual(values = color_vals) +
  #theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + 
  labs(y = 'n. of studies')
p2e
```

