---
title: "fig5"
author: "Daniel Chawla"
date: "3/18/2022"
output: md_document
---

Figure 5: Nearly all infection signatures are cross-reactive with unin- tended infections or non-infectious conditions

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# function that filters AUROCs to those calculated using at least frac_threshold (fraction of) signature genes
filterGeneFraction <- function(input_df, frac_threshold){
  input_df %>% 
    mutate(frac_pos = ifelse(is.na(frac_pos), 1, frac_pos)) %>%
    mutate(frac_neg = ifelse(is.na(frac_neg), 1, frac_pos)) %>% 
    filter(frac_pos >= frac_threshold, frac_neg >= frac_threshold) %>%
    return()
}

input_XR_file <- '~/Dropbox/big_eval_heatmap_list_specificity_021322_healthy_sbj_ts.RDS'
```

## Figure 5A & C cross-reactivity of viral and bacterial signatures

Cross-reactivity is evaluated using `analysis/calc_robustness_crossreactivity.R`
```{r}
library(ggplot2)
library(dplyr)
library(openxlsx)
library(ggthemes)
library(stringr)
library(scales)
library(ggh4x)
library(ggpubr)
library(ggradar)
library(ggridges)
library(grid)


source('../analysis/helperFunctions.R')
source('../analysis/plot_palette.R')

# function to order data according to specificity stratified by input column
getOrder <- function(input_data, col_name){
  col_name <- enquo(col_name)
  order <- input_data %>%
    group_by(!!col_name) %>%
    summarize('S' = sum(Specific.Logical)) %>%
    arrange(desc(S)) %>%
    pull(!!col_name)
  return(order)
}


# load signatures
signatures_file <- '../tables/supp_1.xlsx'
signatures <- read.xlsx(signatures_file)

# load cross-reactivity calculations
plot_file <- input_XR_file
out <- readRDS(plot_file)

plot_dat <- bind_rows(out) %>%
  filter(Type != 'Null') %>% # remove permuted labels
  mutate('Type' = factor(Type, levels = c('Discovery', 'Validation', 'New', 'Null'), ordered = T)) %>%
  mutate(Comparison = str_extract(pattern = 'VxB|V|B', string = Signature)) # identify signature comparison

N_threshold = 5 # closed bound
colors <- createColorScale()

input_tmp_plot_dat <- plot_dat %>%
  filter(N >= N_threshold, Type != 'Null') %>%
  mutate(Signature = as.character(Signature))

sig_order <- rev(c(paste0('V', 1:11), paste0('B', 1:7),  paste0('V/B', 1:7)))
study_plot_dat <- input_tmp_plot_dat %>%
  mutate(Signature2 = factor(Signature, levels = sig_order, ordered = T)) %>%
  filter(!grepl(pattern = 'I', x = Signature)) 

color_vals <- c("#79d2f0", "#ffa31f", "#1fe58c", "gray50")
names(color_vals) <- c('V', 'B', 'V/B', 'non-specific') # swap bacterial and viral for this case 
specificity_threshold = 0.6
specificity_plot_dat <- study_plot_dat %>%
  filter(!Signature %in% c('B1', 'V3')) %>% 
  filter(!grepl(pattern = 'V/B', x = Signature)) %>%
  group_by(Signature) %>%
  mutate('med' = median(Scores)) %>% 
  mutate(Color = as.character(Comparison)) %>%
  mutate(Color = ifelse(med <= specificity_threshold, Color, 'non-specific')) %>% 
  mutate(Signature = factor(Signature, levels = sig_order, ordered = T)) %>%
  mutate(Comparison = factor(Comparison, levels = c('V', 'B'), ordered = T))


p3 <- ggplot(specificity_plot_dat, aes(x = Scores, y = Signature, fill = Color, color = Color)) + 
  geom_density_ridges(quantile_lines = F, alpha = 0.9) + 
  #geom_vline(xintercept = specificity_threshold, linetype = 'dashed', color = 'blue') + 
  #geom_vline(xintercept = c(0.5), linetype = 'dashed', color = 'black') + 
  scale_fill_manual(values = color_vals) + 
  scale_color_manual(values = color_vals) + 
  labs(x = 'AUC') + 
  facet_grid(Comparison ~ ., scales = 'free', space = 'free', switch = 'y') +  
  #xlim(c(0,1)) + 
  theme_clean + 
  theme(legend.position = 'none', strip.background = element_rect(fill = 'white')) + 
  scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.2), limits = c(0,1)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

p3
```


## 5B — viral specificity by bacterial gram stain
Gram stain information is available in `tables/bact_gram_stains.xlsx`

```{r}

# load gram stain information
gram_stains <- read.xlsx('../tables/bact_gram_stains.xlsx') %>%
  mutate(Include = as.logical(Include)) %>%
  dplyr::select(Gram.stain, Study, Include, Age.Group, Tissue) %>%
  mutate(Gram.stain = gsub(pattern = 'acid-fast', replacement = 'Acid-fast', x = Gram.stain)) %>%
  mutate(Gram.stain = gsub(pattern = 'Pos', replacement = 'Gram positive', x = Gram.stain)) %>%
  mutate(Gram.stain = gsub(pattern = 'Neg', replacement = 'Gram negative', x = Gram.stain)) %>%
  mutate(Gram.stain = gsub(pattern = 'Mixed', replacement = 'Multiple gram types', x = Gram.stain)) %>%
  mutate(Gram.stain = gsub(pattern = 'not reported', replacement = 'Unreported', x = Gram.stain)) %>%
  mutate(Gram.stain = factor(Gram.stain, levels = c('Acid-fast', 'Gram positive', 'Multiple gram types', 'Gram negative', 'Unreported'), ordered = T))

# gse77528 is parasitic
# gse73464 is non-infectious
accessions_to_remove <- c('GSE77528', 'GSE73464')
plot_dat <- plot_dat %>%
  filter(!grepl(pattern = paste(accessions_to_remove, collapse = '|'), x = Study))

input_tmp_plot_dat <- plot_dat %>%
  filter(N >= N_threshold, Type != 'Null', Comparison == 'V') %>%
  mutate(Signature = as.character(Signature))

signature_labels <- unique(input_tmp_plot_dat$Signature)

# join plotting data with gram stain information
study_plot_dat <- input_tmp_plot_dat %>%
   mutate(Train_Label = ifelse(Type == 'Discovery', '○', '')) %>%
   left_join(gram_stains) %>%
   mutate('Signature' = factor(Signature, levels = signature_labels, ordered = T)) #%>%
   #mutate('Study' = factor(Study, levels = order_dat$Study), ordered = T)

gram_plot_dat <- study_plot_dat %>%
  group_by(Gram.stain, Signature) %>%
    summarize(N = n(), AUC = median(Scores))

# plotting
shape_vals <- c(22, 21, 24)
names(shape_vals) <- c('Acid-fast', 'Gram positive', 'Gram negative')
color_vals <- c('#79d2f0', 'gray50')
names(color_vals) <- c('NXR', 'XR')
p5c <- ggplot(gram_plot_dat %>%
                filter(!Signature %in% c('V3')) %>% 
                mutate(Signature = factor(Signature, levels = paste0('V', 11:1), ordered = T)) %>% 
                filter(grepl(pattern = 'positive|negative|fast', x = Gram.stain)) %>%
                mutate(XR = ifelse(AUC <= specificity_threshold, 'NXR', 'XR')), 
              aes(x = Signature, y = AUC, shape = Gram.stain, group = Gram.stain, fill = XR, color = XR)) + 
  geom_line(size = 2, color = 'gray80') + 
  geom_point(size = 4) +
  scale_y_continuous(limits = c(0,1), breaks = seq(to = 1, from = 0, by = 0.2)) + 
  theme_clean + 
  scale_shape_manual(values = shape_vals, name = 'bacterial class') + 
  scale_fill_manual(values = color_vals, name = '') + 
  scale_color_manual(values = color_vals, name = '') + 
  theme(legend.position = 'top') + 
  coord_flip() + 
  guides(fill = 'none')

p5c2 <- p5c + theme(legend.text = element_text(size = 10), legend.position = 'bottom')
p5c2

```

## 5D-F: Cross-reactivity with parasitic infections, obesity, and aging

Cross-reactivity is evaluated using `analysis/calc_robustness_crossreactivity.R`

```{r, include = F, echo = F}
parasite_cross_reactivity_file <- '~/Documents/darpa-manuscript-data/parasite_eval_heatmap.RDS'
aging_obesity_file <- '~/Documents/darpa-manuscript-data/figures/aging_obesity_heatmap_012522.RDS'


aging_list <- readRDS('~/Documents/darpa-manuscript-data/data_list_final/aging_list.RDS')
obesity_list <- readRDS('~/Documents/darpa-manuscript-data/data_list_final/obesity_list.RDS')

exposure_df <- lapply(c(aging_list, obesity_list), function(x){
  data.frame('Study' = x$formattedName, 'Exposure' = setdiff(unique(x$pheno$Standardized.Exposure), 'Healthy')) %>%
    return()
}) %>%
  bind_rows()
```


```{r}
parasite_df <- readRDS(parasite_cross_reactivity_file)
sigs_to_omit <- c('V3', 'B1')

specificity_threshold <- 0.6

plot_dat_para <- parasite_df %>%
  dplyr::filter(Sig.Type != 'Influenza') %>%
  dplyr::filter(!Signature %in% sigs_to_omit) %>%
  mutate(Sig.Type = gsub(Sig.Type, pattern = 'x', replacement = '/')) %>%
  mutate(Sig.Type = gsub(Sig.Type, pattern = 'Virus', replacement = 'V')) %>%
  mutate(Sig.Type = gsub(Sig.Type, pattern = 'Bacteria', replacement = 'B')) %>%
  mutate(Sig.Type = factor(Sig.Type, levels = c('V', 'B', 'V/B'))) %>% 
  group_by(Signature) %>%
  mutate(med = median(Scores)) %>% 
  mutate(color = ifelse(med > specificity_threshold, 'not specific', as.character(Sig.Type))) %>%
  ungroup() %>%
  mutate(Signature = gsub(pattern = 'x', replacement = '/', x = Signature)) %>% 
  mutate(Signature = factor(Signature, levels = sig_order, ordered = T))

color_vals <- c("#79d2f0", "#ffa31f", "#1fe58c", "gray50")
names(color_vals) <- c('V', 'B', 'V/B', 'not specific')
color_vals_para <- color_vals
#color_vals_para[c('V', 'B', 'V/B')] <- c(rep('gold1', 3))
f5b <- ggplot(plot_dat_para %>% mutate('Exposure' = 'Parasite'), 
       aes(x = Signature, y = Scores, fill = color)) + 
  geom_boxplot(outlier.shape = NA) +
  #geom_boxplot() +
  #geom_density_ridges() + 
  #geom_point(data = outlier_df, color = 'black', pch = 4) + 
  theme_clean + 
  #geom_point(data = df %>% 
  #             filter(grepl(pattern = 'GSE1124', x = Study), Sig.Type == 'VxB'), 
  #           aes(color = 'red')) + 
  #geom_line(aes(group = Study), alpha = 0.3) + 
   facet_grid(Exposure ~ Sig.Type, space = 'free', scales = 'free', switch = 'y') + 
  # theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_fill_manual(values = color_vals_para) + 
   #geom_hline(yintercept = c(0.6), linetype = 'dashed', color = 'blue') +
   #geom_hline(yintercept = 0.5, linetype = 'dashed', color = 'black') +
   #labs(x = 'AUROC') + 
  theme(strip.background = element_rect(fill = 'white')) + 
  scale_y_continuous(breaks = seq(from = 0, to = 1, by = 0.2), limits = c(0,1)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), legend.position = 'none') + 
  labs(y = 'AUROC')
f5b  


## aging and obesity next
# read results
aging_obesity_results <- readRDS(aging_obesity_file) 

# format for plotting
plot_dat <- bind_rows(aging_obesity_results) %>%
  filter(!grepl(pattern = '^I', x = Signature))
plot_dat[grepl(pattern = 'GSE65219', x = plot_dat$Study) & grepl(pattern = 'Sampson', x = plot_dat$Signature), 'Type'] <- 'Discovery'

plot_dat <- plot_dat %>%
  mutate('Type' = factor(Type, levels = c('Discovery', 'Validation', 'New', 'Null'), ordered = T)) %>%
  mutate('Comparison' = str_extract(pattern = 'VxB|V|B', string = Signature))  %>%
  arrange(Comparison) %>%
  mutate(Signature = factor(Signature, levels = unique(Signature), ordered = T))

# add exposure labels (aging or obesity)
plot_dat <- plot_dat %>%
  left_join(exposure_df)

N_threshold = 0
#colors <- createColorScale()

# facet by pathogen and cluster rows/columns
plot_list <- list()
head(plot_dat)
table(plot_dat$Exposure.Class)
exposures <- unique(plot_dat$Comparison)

pattern_vals <- c('none', 'stripe')
names(pattern_vals) <- c('Aging', 'Obesity')
f5efg_list <- list()
plot_dat <- plot_dat %>% 
  mutate(Comparison = gsub(pattern = 'VxB', replacement = 'V/B', x = Comparison)) %>%
  mutate(Comparison = factor(Comparison, levels = c('V', 'B', 'V/B'), ordered = T))


#color_vals_noninf <- color_vals_para[2:4]
#names(color_vals_noninf) <- c('Aging', 'Obesity', 'not specific')
f5efg <- ggplot(plot_dat %>%
                  filter(!Signature %in% sigs_to_omit) %>%
                  filter(!is.na(Comparison)) %>% 
                  #filter(Exposure == 'Aging') %>% 
                  mutate(Exposure = factor(Exposure, levels = c('Obesity', 'Aging'), ordered = T)) %>% 
                  group_by(Exposure, Signature) %>%
                  mutate('med' = median(Scores)) %>% 
                  mutate('specific' = ifelse(med <= specificity_threshold, as.character(Exposure), 'not specific')),
                aes(x = Signature, y = Scores, fill = Comparison)) + 
  geom_boxplot() + 
  #scale_pattern_manual(values = pattern_vals) + 
  facet_grid(Exposure ~ Comparison, scales = 'free', space = 'free') +
  #facet_grid(. ~ Comparison, scales = 'free', space = 'free') +
  #ylim(c(0,1)) + 
  scale_fill_manual(values = color_vals) +
  #geom_hline(yintercept = c(0.6), color = 'blue', linetype = 'dashed') +
  #geom_hline(yintercept = c(0.5), color = 'black', linetype = 'dashed') + 
  theme_clean + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + 
  labs(x = '', y = 'AUROC') + 
  theme(legend.position = 'none', strip.background = element_rect(fill = 'white')) + 
  ylim(c(0,1))
f5efg
```


